Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-28 Response.tex

Thank you for the comment.
The introduction is reduced and many of the paragraphs have been reworded and combined. 
In addition, the extended discussion of the benefits of variational integrators is removed from the introduction.
The mathematical background has been condensed into a single section with separate subsections which discuss some of the mathematical details.
The background section is focused on reviewing the relevant material rather than reintroducing topics.
In the interest of space the reorganized section is omitted in this response. 
In short, the manuscript is completely reorganized and all of the background materials are summarized in~\cref{sec:background}.

@@

Thank you for the comment.
The mathematical background has been reduced and condensed into a single section with multiple subsections.
The discussion of the PCRTBP, invariant manifolds and \Poincare maps are combined into a single subsection.
Furthermore, the variational integrator discussion has been reduced and the background material on the derivation of variational mechanics has been omitted.
Instead, only the specific discussion of the development of the variational integrator is included.

\item
    \begin{itshape}
Section 4 ends on page 14 with essentially a thesis statement ("In this paper, we...").  Strive to push the boundary between background and new developments to a much earlier point in the paper; it shouldn't happen so late that you need to remind the audience what the paper is for.
\end{itshape}

Thank you for the comment. 
The introductory detail has been reduced and combined into a single section. 
As a result, the third section begins with the new contributions which describe the optimal control formulation of the reachability set.

\item 
    \begin{itshape}
A missing relevant reference: "On target for Venus a set oriented computation of energy efficient low thrust trajectories" (Dellnitz, Junge, Post, and Thiere, 2006)
\end{itshape}

@@

\begin{itemize}
\item[-] Ober-Blöbaum, Sina, Oliver Junge, and Jerrold E. Marsden. "Discrete mechanics and optimal control: an analysis." ESAIM: Control, Optimisation and Calculus of Variations 17.2 (2011): 322-352.
\end{itemize}

In the revised manuscript, all of the comments for the variational integrator are re-examined and reorganized to Section 2.2, as also discussed in the response to the first reviewer.

\item
    \begin{itshape}
Apart from these claims, I question the need for a variational integrator in the first place. The example in section 3.4 is not convincing because there is no apples-to-apples basis for the comparison; a variable-step integrator with unspecified tolerance is compared to a fixed-step integrator with an arbitrary step size.  Integrator comparisons should consider error in proportion to compute cost, which can be tuned by step size or tolerance.  Also, fixed step seems inappropriate for a path through diverse dynamic regimes, and it's not clear that the problem scope calls specifically for energy preservation (especially since the idea is to use low thrust to mitigate some time cost of exploiting the problem structure).  
\end{itshape}

In the revised manuscript, the computational properties of the variational integrator is explicitly compared with a general purpose integrator in~\Cref{fig:compare_energy}. According to the reviewers suggestion, the error in energy conservation is illustrated for several computational costs. It is shown that the variational integrator exhibits lower conservation error  against a Runge-Kutta method for the similar level of CPU time.

@@

    \node[below right] (x1) at (0,3,2) {\(\vecbf{q}_1\)};
    \filldraw (0,3,2) circle (3pt);

    \tdplotdrawarc[orbit hidden]{(center)}{2}{90}{190}{}{};
    \tdplotdrawarc[orbit,<-]{(center)}{2}{-170}{90}{}{};

    \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{199}{}{};
    \tdplotdrawarc[orbit,<-]{(center)}{3}{-161}{90}{}{};

        \end{tikzpicture}
    \end{scaletikzpicturetowidth}
    \caption{Diagram of the \Poincare map: Periodic orbits will appear as fixed points on the \Poincare section \( \Sigma \). Stability of periodic orbits is clearly visible on the section as successive intersections approach or depart a fixed point.}
\end{figure}

\begin{figure}[H]
        \centering 
        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparison: A stable trajectory is used to test the variational and conventional integrators. 
        The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.}
        [0.5\textwidth]{\includegraphics[width=0.2\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
        \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}}
        [0.5\textwidth]{\includegraphics[width=0.2\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
\end{figure}

@@

        \tdplotdrawarc[tdplot_rotated_coords,grid]{(x0)}{0.5}{-45}{90}{above}{\(\theta_d\)};

        % draw terminal state on reachability set
        \node[tdplot_rotated_coords,label=above:\(\vecbf{x}_f\)] (xf) at ($ (x0)+(-45:1.5) $) {};
        \filldraw (xf) circle (3pt);

        \node[tdplot_rotated_coords,label=below:\(J\)] at (xf) {};

        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
        % place
    \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Reachability set on a \Poincare section: Pictorial representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
            The terminal state, \( x_n\), is the intersection without any control input. 
            Adding a control input allows for the terminal state, \( x_f \), to be displaced by some distance/cost \( J \) as measured on the section.
            We parameterize a specific direction on the section with the angles \( \theta_d\) and seek to maximize the distance between \( x_f \) and \( x_n \).
            Computation of the maximum distance, or reachability, for a variety angles gives a discrete approximation of the reachability set.}
\end{figure}

@@

                    % draw segement dividers
                    \draw [dashed,thick] (t1.center) -- (x1m.center);
                    \draw [dashed,thick] (t2.center) -- (x2m.center);
                    \draw [dashed,thick] (t3.center) -- (x3m.center);
                    \draw [dashed,thick] (tn.center) -- (xnm.center);

                    % draw markers at each subsegment

                    % draw the subsegments
                    \draw [->] (x0.center) to [bend left=30] (x1minus.center);
                    \draw [->] (x1plus.center) to [bend left=30] (x2minus.center);
                    % dotted lines here
                    \draw[decorate sep={1mm}{5mm},fill] (x2plus.east) to [bend left=10] (x3minus.west);

                    \draw [->] (x3plus.center) to [bend left=30] (xnminus.center);
                \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Schematic diagram of the multiple shooting method:
        The complete trajectory is split into a number of sub-segments, and additional interior constraints are included to ensure state and costate continuity.
    Splitting the optimal trajectory into short segments reduces the sensitivity of terminal states to variations of the initial states.}
\end{figure}

\begin{figure}[H]
   \centering
   \includegraphics[width=0.2\textwidth]{moon_orbit} % requires the graphicx package
   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstrating the initial and target orbit.
   Without the low-thrust propulsion, the spacecraft is constrained to the initial periodic orbit. 
   We determine the reachability set to find a transfer trajectory to the target orbit about the moon.}
\end{figure}

@@

\begin{figure}[H]
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{manifold_trajectory} 
                \caption{Position space view of invariant manifold} \label{fig:manifold_trajectory} 
        \end{subfigure}~ %
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{manifold_poincare} 
                \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} \label{fig:manifold_poincare} 
        \end{subfigure} 
        \caption{Invariant manifold transfer: An example transfer using the invariant manifolds is shown in both the position and \Poincare spaces.
        The control free transfer from the initial periodic orbit to the target orbit result in a long time of flight. 
    In addition, the manifold only intersects the target orbit on the descending or far side of the moon.}
\end{figure}
\begin{figure}[H]
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{reach_trajectory} 
                \caption{Position space view of reachability set}  
        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{poincare_compare} 
                \caption{\Poincare section view of reachability set}  
        \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)

        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{reach_transfer} 
                \caption{Transfer trajectory selected from reachability set viewed in the position space}  
        \end{subfigure}~ 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{control_input_l1} 
                \caption{Control input for the selected transfer trajectory}  
        \end{subfigure}~
        \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
        The reachability set is shown in the upper two figures in both the position and \Poincare space.
    From this reachability set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
The optimal control to achieve this transfer is shown in the lower right figure.} 
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_trajectory_zoom.pdf} 
        \caption{Stage 1 Trajectory} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_poincare.pdf} 
        \caption{Stage 1 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_trajectory_zoom.pdf} 
        \caption{Stage 2 Trajectory~} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_poincare.pdf} 
        \caption{Stage 2 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_trajectory_zoom.pdf} 
        \caption{Stage 3 Trajectory} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_poincare.pdf} 
        \caption{Stage 3 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
        \caption{Stage 4 Trajectory~} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage4_poincare.pdf} 
        \caption{Stage 4 \Poincare~section } 
    \end{subfigure}   
    \caption{Stage 1-4 reachability sets: The first four reachability sets visualized in both the position (left) and \Poincare space (right).
        The minimum trajectories from the preceding stages are shown in red, while the next stage is shown in blue.
    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
        \caption{Stage 5 Trajectory~} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_poincare.pdf} 
        \caption{Stage 5 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_trajectory_zoom.pdf} 
        \caption{Stage 6 Trajectory} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_poincare.pdf} 
        \caption{Stage 6 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_trajectory_zoom.pdf} 
        \caption{Stage 7 Trajectory} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_poincare.pdf} 
        \caption{Stage 7 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_trajectory_zoom.pdf} 
        \caption{Stage 8 Trajectory} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_poincare.pdf} 
        \caption{Stage 8 \Poincare~section } 
    \end{subfigure}   
    \caption{Stage 5-8 reachability sets: The last four reachability sets visualized in both the position (left) and \Poincare space (right).
        The minimum trajectories from the preceding stages are shown in red, while the next stage is shown in blue.
    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit}
\end{figure}
\end{correction}

@@

\item
    \begin{itshape}
Figure 4 does not show what I think you intend it to show; it looks like two periodic circular orbits.
\end{itshape}

Thank you for the comment, however the figure is drawn to show a pictorial representation of a \Poincare section.
The periodic orbits intersect the section at fixed points, as a result it becomes simple to visualize and determine these fixed solutions by searching for fixed points on the section
Conversely, stable/unstable trajectories will also appear on the section with a distinctive signature, and by tracking successive intersections of the trajectory with the \Poincare section.

@@

        \tdplotdrawarc[tdplot_rotated_coords,grid]{(x0)}{0.5}{-45}{90}{above}{\(\theta_d\)};

        % draw terminal state on reachability set
        \node[tdplot_rotated_coords,label=above:\(\vecbf{x}_f\)] (xf) at ($ (x0)+(-45:1.5) $) {};
        \filldraw (xf) circle (3pt);

        \node[tdplot_rotated_coords,label=below:\(J\)] at (xf) {};

        \tdplotdrawarc[tdplot_rotated_coords,reachability]{(x0)}{1.5}{0}{360}{}{};
        % place
    \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Reachability set on a \Poincare section: Pictorial representation of the reachability set (blue circle) on the \Poincare section, \(\Sigma\). 
        }
\end{figure}
\end{correction}

@@

\begin{correction}
The numerical examples presented in this section satisfy necessary conditions for local optimality, and obtaining a globally optimal solution is considered beyond the scope of this paper.
\end{correction}

\item
    \begin{itshape}
How damaging is the assumption of a fixed travel time?  And how about the specification of a specific departure point on the initial orbit?
\end{itshape}

The fixed travel time is primarily used at this stage to reduce the complexity of the optimal control formulation. 
Furthermore, the reachability set is defined in terms of a fixed terminal time.
A variable terminal travel time would add an additional dimension, time, to the reachable space and add additional complexity to the reachability set visualization.
The fixed travel time does not adversely affect the optimality of the computed trajectories on the reachable set, as the reachable set is defined in terms of a fixed travel time.

The choice of starting from a fixed point on the initial periodic orbit is again mostly used for convenience. 
In this case of the periodic orbit the departing trajectories are all along the unstable invariant manifold associated with the periodic orbit. 
The uncontrolled trajectories associated with the invariant manifolds serve as the ``center'' of each reachability set computation.
The low-thrust propulsion system is then used to enlarge the distance between the uncontrolled and controlled states. 
Starting from a different location on the periodic orbit will not create a qualitatively different situation as the entire invariant manifold also intersects the \Poincare section.
However, different locations along the periodic orbit may have vastly different travel times before intersecting the \Poincare section.
This will require modification of the optimal control problem, with a larger terminal time, but will result in a similar behavior regardless of the location on the initial periodic orbit.

@@

\item
    \begin{itshape}
In Fig 10b, it is hard to see, but appears that the low-thrust-reachable set is actually *centered* on part of the target set.  Is this related to the existence of the uncontrolled transfer opportunity?  There must be some more clear way to show how the low thrust set expands upon the zero thrust set.  Perhaps beginning with the time horizon of the zero-thrust opportunity, and successively lowering it, would help.  
\end{itshape}

### 2017-07-28 Response.tex

The reachability set is closely spaced in terms of both the position space as well as on the \Poincare section.
This is due to the fact that there is a very low-thrust control input on the spacecraft.
As a result, the trajectories associated with the reachability set are all closely spaced, but more widely separated from the terminal state without any control input.
In the figure, there is a fixed point labeled \( x_n \) which is the terminal state without any control input, which will also lie on the periodic orbit.
The reachable set, in red, is widely separated from the uncontrolled state due the effect of the low thrust propulsion, however the subsequent spread of the entire reachability set is much more constrained.
The low thrust expands the reachability set from a fixed point to the red elliptical region. 
This is also explained in the manuscript with a similar description of the figure.

@@

Conducting a single reachability analysis becomes more complicated as the larger time of flight increases the sensitivity of the terminal states to variations in the initial costates. 
Furthermore, the larger time of flight creates a much larger and more chaotic reachable set for the analysis.
Through this work it was found that shorter trajectories tended to have more stable and consistent solutions.

\item
    \begin{itshape}
In general, please devote more depth to probing the strengths, weaknesses, and fundamental properties of the method.  Considering extreme cases can help with this.
\end{itshape}

The authors believe that the given two nontrivial numerical examples illustrate the efficacy of the proposed approach in designing low-thrust maneuvers through reachability set analysis on the \Poincare section. 
The examples demonstrate the ability to generate transfer trajectories in the PCRTBP and can be easily extended to non-planar transfers~\cite{kulumani2016d}.
The proposed approach provides for a systematic design process for low-thrust orbital transfers, in contrast to other methods which must rely entirely on numerical optimization to determine trajectories.
This design process leverages the reachability set on a \Poincare section to greatly reduce the computational complexity of an unmotivated search of the reachable set.
Through the use of the reachability set, our proposed approach provides a methodical process to determine transfers, using single or multiple iterations.

There are several drawbacks of this approach.
First, it relies heavily on the PCRTBP, which limits the applicability to planar trajectories.
Furthermore, this reduces it's benefit to operational missions which are often non-planar, e.g. LEO to GTO missions.
Second, the method does not consider the effect of variable mass on the system dynamics. 
This is a focus of future work to incorporate the effect of variable mass on the optimal control problem.
Finally, the variational integrator provides for a bounded energy behavior but does not ensure exact energy over the simulation. 
As a result, there is still error in the variational integrator with respect to energy deviation.

The manuscript has been completely reorganized to emphasize the contributions and clearly separate the background material. 
In particular,~\cref{sec:simulation} includes additional paragraphs and figures for the geostationary to periodic orbit transfer.
These additional figures demonstrate the ability to chain together multiple iterations of the reachability set computation to generate a nontrivial transfer trajectory.
Combined together, the numerical examples demonstrate the ability to depart from the vicinity of the Earth to the vicinity of the Moon using our proposed method of coupling low-thrust maneuvers with reachability set analysis.

@@

In addition, the conclusion has been modified to better demonstrate the strengths and weaknesses of our approach as well as avenues for future research.
The conclusion is copied below.

\begin{correction}

In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution.
Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.
